Staged Abstract Interpreters

### 2019-04-05 paper/60_cases.tex
paper

In this section, to demonstrate the SAI approach is widely applicable, we extend
and refactor the staged abstract interpreter we presented in Section \ref{sai}
to various control-flow analysis (CFA). We first discuss calling-context
sensitive analysis for staged analyzers; then we will see how to implement
store-widening with staging, which is a common optimization in CFA; finally, we
integrate abstract garbage collection with staging, \todo{where the addresses to
  be collected can be computed at the staging time}.

@@

% \subsubsection{Mixed Sensitivity}
% using different k?

One of the challenges of modern static analysis is program usually depends on
large libraries programs~\cite{toman_et_al:LIPIcs:2017:7121}. Can we analyze
programs and libraries separately and reuse the result without losing precision?
Then we can reduce part of the overhead of repeatedly analyzing libraries for
different programs. Indeed, some static analyzers compute summary for a function
or a module, which can be reused later, however they are mostly too conservative
or unsound, which both lead to imprecision.

The specialization of abstract interpreter provides a chance to obtain such
partial analysis result in a mechanized way, but still keeps the analysis sound.
As we see when compiling the closures, we can specialize the abstract
interpreter with respect the body expression of the lambda term without knowing
the actual argument. The programs with some unknown variables are open programs,
which is exactly the case if we want to analyze programs in a modular way.

For a concrete analysis, for example, $k$-CFA ($k$ > 0) is naturally a
whole-program analysis, because it is inter-procedural and needs the last $k$
calling contexts to distinguish different call sites, where the calling contexts
are dynamic values. However, it is possible to analyze programs (libraries)
separately through specializing an abstract interpreter that generates the
analysis as the next-stage program and leave the unavailable programs and
calling contexts as dynamic parameters, and then install these contexts when we
have the whole program.

>>>>

In this section, to demonstrate the SAI approach is widely applicable, we extend
and refactor the staged abstract interpreter we presented in Section \ref{sai}
to various control-flow analysis (CFA). We first discuss calling-context
sensitive analysis for staged analyzers by tweaking allocation strategy; then we
will see how to implement store-widening with staged monads, which is a common
optimization in CFA; finally, we integrate abstract garbage collection with
staging.

@@

% using different k?

One of the challenges of modern static analysis is program usually depends on
large libraries programs~\cite{toman_et_al:LIPIcs:2017:7121}. If we can analyze
programs and libraries separately and reuse the result without losing precision,
then we can reduce the overhead of repeatedly analyzing libraries.
Some static analyzers compute a summary for a function or a module, which
can be reused later, however they are mostly too conservative or unsound, which
both lead to imprecision. In this section, we show meta-programming can be used
to create sound summaries in the code generation from, such generated code can
be resued later. The summary contains information that not depend on the
function argument.

As we see when compiling a higher-order function (closure), we specialize the
abstract interpreter to the body expression of that lambda term under some
environment, without knowing the actual arguments to the function. We may reuse
the infrastratures and refactor the @close@ function mentioned in Section
\ref{sai} to achieve modular analysis. The @close@ function takes a lambda term
@f@ and an environment, produces a compiled closure, which can be freely used at
the next stage.

\begin{lstlisting}
  def close(ev: Eval => Ans)(f: Expr, ρ: Rep[Env]): Rep[CompiledClo] = { ...}
\end{lstlisting}

\paragraph{Partially-Static Context} Under a polyvariant analysis (like
$k$-CFA), such summary generated by specializing static analyzer even could do
more for calling context by partially-static data optimization. For example,
consider that we have a library consists of three functions @f@, @g@ and @h@,
where @f@ calls @g@ and @h@ internally and returns @y@ fianlly (we extend the
language to have multiple arguments). @g@ and @h@ do not depend on @f@. Now we
would like to modularlly specialize these functions, starting from @f@.

\begin{lstlisting}
  def f(a, b, c) = val x = g(a, b); val y = h(x, c);  y
\end{lstlisting}

Since $k$ is bounded to a fixed number, after the analysis running $k$ steps
inside of @f@, the calling context will purely depend on the static structure of
our module being analyzed.
